\documentclass{siamltex}

\advance\oddsidemargin by -0.25in
\advance\evensidemargin by -0.25in
\advance\textwidth by .55in

\usepackage{amsfonts,amssymb,amsmath}
%\usepackage{subfigure}
\usepackage[dvips,pdftex]{graphicx}
\usepackage{subfig}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{url}
\usepackage{calc,array}
\usepackage{booktabs}
\usepackage[usenames,dvipsnames]{color}
\usepackage[linesnumbered,noline,ruled,noend]{}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{rem}[thm]{Remark}
\DeclareMathOperator*{\argmin}{argmin}
\newcommand{\Bcoarse}{\ensuremath{B^{C}}}
\newcommand{\nns}{near null-space}
\DeclareGraphicsRule{.pdftex}{pdf}{*}{}
\usepackage{algorithmic}
\usepackage{algorithm}
\numberwithin{algorithm}{section}  % <--- chapter, section etc. depending on what is required
\def\Boxers#1#2{\noindent$\hphantom{Muv}${\parbox[t]{1.75in}{\it
#1}}{\parbox[t]{2.9in}{#2}} \\[0.7em]}

\title{Level Interface: getting at level data in MueMat/MueLu}

\author{J.~Hu
\and    J.~Gaidamour
\and    R.~Tuminaro
\and    T.~Wiesner}

% For making comments in the document.
\newcommand{\JJH}[1]{\textcolor{red}{JJH: #1}}
\newcommand{\JG}[1]{\textcolor{cyan}{JG: #1}}
\newcommand{\RST}[1]{\textcolor{Blue}{RST: #1}}
\newcommand{\TW}[1]{\textcolor{RoyalBlue}{TW: #1}}


\begin{document}
\maketitle


\section{Test Cases} \label{test cases}
Some cases to consider (note: auxiliary matrix and Emin(Emin) examples are interesting but would not be used heavily).
\vskip .1in
\begin{enumerate}
\item SA with coalesced $A$ for aggregation.\\

\item SA with filtered then coalesced $A$ for aggregation \& filtered $A$ for
   prolongator smoother.\\

\item SA with user-defined auxiliary graph for aggregation \& a decoalesced
   version to define matrix for prolongator smoother.\\

\item SA with user-defined auxiliary graph for aggregation \&
   prolongator smoother matrix is a filtered $A$ where pattern
   matches decoalesced version of auxiliary matrix.
   \newcounter{enumi_saved}
    \setcounter{enumi_saved}{\value{enumi}}
\end{enumerate}
\vskip .1in
%Emin algorithms are trickier. The main additional things to play with include:
%\vskip .1in
%\Boxers{P0ForPattern}{obvious}
%\Boxers{P0ForKrylovInitialGuess}{obvious}
%\Boxers{P0ForAPpattern}{obvious}
%\Boxers{AForAPpattern}{obvious}
%\Boxers{AForEnergy}{obvious}
%\Boxers{MForKrylov}{might want to consider preconditioning the Krylov 
%minimizer with arbitrary block diagonals.}
%\Boxers{FNull,CNull,Constraints}{Only MakeConstraints need FNull and CNull.
%                              While FNull/CNull might be defined in different
%                              ways, I'm not sure that we would ever have a
%                              valid case where we would want more than one
%                              type of null space on the same level. We might
%                              someday want to consider constraints other than
%                              P CNull = FNull.  I'm thinking mostly of
%                              something where we only care about adding
%                              constraints locally to enforce local features
%                              like preserving a discontinuity.}
\begin{enumerate}
      \setcounter{enumi}{\value{enumi_saved}}
\item Standard Emin using an initial guess of $Ptent$ (obtained by aggregating
   coalesced A).\\
\item Badri-like case where two user-provided matrices are each filtered
      to provide different matrices for Emin: one is Krylov
      energy \& the other for use by APpattern.\\
\item Emin using 1-step SA as initial guess.\\

\item Emin using Emin as initial guess where each computes a Ptent and Coarse Null space \& both are to be saved for later reuse.\\

\item Convert an SA factory so that it can be used for restriction.\\

\item A test case for an F/C-style sparsity pattern (e.g,
      F/C pattern generates a P0)?

\end{enumerate}
\section{Function Pointer Interface}
\subsection{LevelInterface class}
A simple class with only two things: static functions and data member 
function pointers The basic idea is that the function pointers are used in
specific factories to access data. They effectively define the factory's
interface to Level.  They are set to one of the particular static functions. 
Developers add new static functions as needed and perhaps new function 
pointers as well.  There is no attempt to group functions into subclasses
or to hide some functions from some factories.  The static functions
are extremely basic and are not meant to hide complex logic. They are
strictly for data access.  A sample matlab-style class is shown below
\vskip .1in
\begin{verbatim}
classdef LevelInterface < CopiableHandle

  properties (Access = public)

    % function pointers (null or set to a static function) defining interface.
    
   P0ForPattern            = 
   P0ForKrylovInitialGuess = @LevelInterface.GetPtent
   P0ForAPpattern          = @LevelInterface.GetPtent
   AForAPpattern           = @LevelInterface.GetA
   AForEnergy              = @LevelInterface.GetA
   GraphForCoarsening      = @LevelInterface.GetGraphOfA
   NullForPtent            = @LevelInterface.GetNull
   NullForConstraints      = @LevelInterface.GetNull

   SetFilteredA            = @LevelInterface.SetFilteredA
   SetGraph                = @LevelInterface.SetGraph
   SetCoalescedA           = @LevelInterface.SetCoalescedA
   SetNull                 = @LevelInterface.SetNull
   MForKrylov
   Constraints
   CPoints

  end %properties

  methods (Static = true)

    % static functions for use by the above function ptrs

    function [A] = GetA(Level)
        A = Level.GetA();               end 
    function [A] = GetATrans(Level)
        A = Level.CheckOut("ATrans");   end
    function [A] = GetFilteredA(Level)  
        A = Level.CheckOut("Afiltered");end
    function [A] = GetFilteredAAux(Level)  
        A = Level.CheckOut("AfiltAux"); end
    function [A] = GetCoalescedA(Level)
        A = Level.CheckOut("ACoalesce");end
    function [P] = GetP(Level)        
        P = Level.GetP();               end
    function [P] = GetPtent(Level)    
        P = Level.CheckOut("Ptent");    end
    function [P] = GetPAux(Level) 
        P = Level.CheckOut("PAux");     end 
    function [G] = GetGraphOfA(Level) 
        G = Level.CheckOut("Graph");    end
    function [N] = GetNull(Level)     
        N = Level.CheckOut("Null");     end
    function [N] = GetNullAux(Level)     
        N = Level.CheckOut("NullAux");  end
    function [ ] = SetA(Level, A)          
        Level.SetA();                   end
    function [ ] = SetFilteredA(Level,A)    
        Level.Save("Afiltered",A);      end
    function [ ] = SetFilteredAAux(Level,A)    
        Level.Save("AfiltAux",A);       end
    function [ ] = SetCoalescedA(Level,A)  
        Level.Save("ACoalesce",A);      end
    function [ ] = SetNull(Level,Null)          
        Level.Save("Null",Null);        end
    function [ ] = SetP(Level,P)          
        Level.SetP();                   end 
    function [ ] = SetPtent(Level,P)      
        Level.Save("Ptent",P);          end
    function [ ] = SetPAux(Level,P)   
        Level.Save("PAux",P);           end
    function [ ] = SetGraphOfA(Level,G)
       Level.Save("Graph",G);           end

  end % methods (Static = true)

end %Operator classdef
\end{verbatim}
\vskip .1in
Note: CheckOut() and Save() are MueLu ideas to be incorporated in MueMat.
To simplify the presentation, data members are assumed public
(things can obviously done via Set/Get). 

\subsection{Factory Access}
One can make arguments in favor of a Level-centric view or a Factory-centric
view. Here, a bit of both is proposed where each factory makes the final 
decision.  In particular, each Level has a field which is of type 
LevelInterface called LevelsView.  Inheritance would be highly discouraged
for this LevelsView.  When users create a level, they can do things like
\vskip .1in
\begin{verbatim}
    FineLevel.LevelsView.AForAPpattern = @GetFilteredA;
    FineLevel.LevelsView.AForEnergy    = @GetFilteredA;
\end{verbatim}
\vskip .1in
Again, Set/Get functions can be used to hide things.  When a new coarse level 
is created by Populate(), it copies FineLevel.LevelsView.  In this way, all 
levels have a well-defined interface. If the user wants different behavior 
on a coarser level, he can manually create a coarse level and set its
LevelsView function pointers to whatever he wants.

So far, this looks Level-centric, but we can allow each factory to decide
what it does with the interface. For example, a factory could have a private
member of type LevelInterface called FactorysView.  The factory's build() 
now has two LevelInterfaces: FactorysView and LevelsView.  The build() 
implementer decides how these are merged to create a consistent interface 
called MergedView that is used within the method. Inheritance is permitted in
FactorysView and MergedView.  As a merge example, consider 
\begin{itemize}
\item function pointers in FactorysView are NULL except for a couple 
      where behavior is to be altered in a factory specific way.
\item build() copies LevelsView to MergedView (which is local to 
      build()).  It then copies all nonNull function pointers 
      from FactorysView to MergedView. 
\item MergedView is now used exclusively within build().
\end{itemize}

This is meant as a simple example to give the idea. Factories could
do powerful (but perhaps dangerous things) if they wanted. For example,
suppose that build() not only wants to change how it accesses
data but how other build methods in other factories that it calls access the 
Level. It could save LevelsView, then copy MergedView into the Level structure
(overwriting LevelsView), and finally just before exiting the
build function it could restore the saved copy back in the Level. If
inheritance is discouraged in LevelsView, then only an uninherited version
of the MergedView is stored in the Level.  There are other possibilities. 
Another merge idea (mimicking Jeremie's view of different SmartLayers 
building on each other) is a capability to define a MergedView function 
pointer as a copy of a DIFFERENT LevelsView function pointer.  For example,
SA by definition uses the same matrix to define both energy and the AP pattern.
One way to recognize this in the interface is via
\vskip .1in
\begin{verbatim}
     MergedView.AForAPpattern = LevelsView.AForAPpattern. 
     MergedView.AForEnergy    = LevelsView.AForAPpattern. 
\end{verbatim}
\vskip .1in
That is, MergedView is just a reshuffling of function pointers in 
LevelsView. This effectively defines a mapping from one interface to
a new one. There are also other ambitious possibilities. For example, a 
factory could use the Save/CheckOut mechanism to store additional 
interfaces within a level's data.  So long as another factory knows what 
to look for, it can pull out these other interfaces and use them when
forming its MergedView.  

\subsection{Strengths}
\begin{itemize}
\item Handles important cases.
\item Easy to understand without looking at much documentation.
\item Factory implementers have a lot of freedom while allowing a
      basic mechanism to associate an interface with a level.
\item Could probably be changed easily if we decide later to adopt another
      model (Tobias' Communication Layer or Jeremie's SmartLevel).
\end{itemize}
\subsection{Weakness}
\begin{itemize}
\item Does not logically group or organize access functions.
\item Factories could do dangerous things.
\item \TW{The \textit{LevelInterface} class emulates the old-style \textit{Level} class interface. You may need auxiliary variables in the \textit{InterfaceLevel} class (e.g. \textit{GetPAux}, for test case 7). If we have a highly flexible \textit{Level} class with a hashtable, it doesn't make sense for me to introduce an interface class for the \textit{Level} class that uses predefined auxiliary matrices like \textit{GetPAux}. In my opinion we should try to make use of the \textit{Level} class and not hide it behind an interface like \textit{LevelInterface}.}
\item \TW{not object-oriented $\rightarrow$ port to MueLu?}
\item Bring it on ... I can take it!
\end{itemize}


\subsection{Implementation of Test Cases}
\begin{enumerate}
\item SA with coalesced A for aggregation.
   \begin{verbatim}
   Standard SA case requiring only defaults.
   \end{verbatim}
\item SA with filtered then coalesced $A$ for aggregation \& filtered $A$ for
   prolongator smoother.\\
   \begin{verbatim}
   FineLevel.AForAPpattern = @LevelInterface.GetFilteredA;
   FineLevel.AForEnergy    = @LevelInterface.GetFilteredA;
   \end{verbatim}
\item SA with user-defined auxiliary matrix for aggregation \& a decoalesced
   version to define matrix for prolongator smoother.\\
   \begin{verbatim}
   FineLevel.AForAPpattern      = @LevelInterface.GetFilteredA
   FineLevel.AForEnergy         = @LevelInterface.GetFilteredA

   This strange case can be addressed by a version of CoalesceAndDrop
   which decoalesce graphs. Instead of polluting LevelsView, we make an
   extended version of FactorysView with function pointers
           GetGraphForDecoalescing = @LevelInterface.GetGraphOfA
           SetDecoalescedA         = @LevelInterface.SetFilteredA.
   The user puts the Graph in the right fine level place and provides
   code to project graphs to coarse levels.
   \end{verbatim}
\item SA with user-defined auxiliary graph for aggregation \&
   prolongator smoother matrix is a filtered $A$ where pattern
   matches decoalesced version of auxiliary matrix.
   \begin{verbatim}
   Same as above. Special CoalesceAndDrop decoalesces, but does not
   store decoalesced matrix. Instead it is used to filter A and store 
   result in standard filtered spot(thus, no need for SetDecoalescedA)
   \end{verbatim}
\item Standard Emin using an initial guess of $Ptent$ (obtained by aggregating
   coalesced A).
   \begin{verbatim}
   Just use defaults.
   \end{verbatim}
\item Badri-like case where two user-provided matrices are each filtered
      to provide different matrices for Emin: one is Krylov
      energy \& the other for use by APpattern.
   \begin{verbatim}
   FineLevel.AForAPpattern      = @LevelInterface.GetFilteredA
   FineLevel.AForEnergy         = @LevelInterface.GetFilteredAAux

   The standard CoalesceAndDrop factory is extended so that it can 
   filter two different matrices and store them. Inheritance is
   used to extend FactorsView so that it has functions like
       GetMatrixForFilteredEnergy,  SetMatrixForFilteredEnergy 
       GetMatrixForFilteredPattern, SetMatrixForFilteredPattern
   \end{verbatim}
\item Emin using 1-step SA as initial guess.
   \begin{verbatim}
      SAFactorysView.SetP           = @LevelInterface.SetPAux
      EminFactorysView.P0ForPattern = @LevelInterface.GetPAux
      EminFactorysView.P0ForKrylovInitialGuess= @LevelInterface.GetPAux
   \end{verbatim}
\item Emin using Emin as initial guess where each computes a Ptent and Coarse Null space \& both are to be saved for later reuse.
   \begin{verbatim}
   Create two Emin factories: OuterEmin & InnerEmin. InnerEmin is
   standard. OuterEmin is changed so that it Sets/Gets the Ptent
   and CNull from nonstandard spots.
      OuterEminFactorysView.GetPtent = @LevelInterface.GetPAux
      OuterEminFactorysView.SetPtent = @LevelInterface.SetPAux
      OuterEminFactorysView.GetNull  = @LevelInterface.GetNullAux
      OuterEminFactorysView.SetNull  = @LevelInterface.SetNullAux
   Also modify Needs/CrossFactory string so counters are properly set?
   \end{verbatim}
\item Convert an SA factory so that it can be used for restriction.
   \begin{verbatim}
   Extend SA factory so that it computes/stores A^T, computes/stores
   Rtent^T, and invokes standard build() with
        FactorysView.GetPtent      = @LevelInterace.GetTransRtent
        FactorysView.AForAPpattern = @LevelInterface.GetATrans
        FactoriysView.AForEnergy   = @LevelInterface.GetATrans
        FactorysView.SetPtent      = @LevelInterace.SetTransRtent
   \end{verbatim}
\item A test case for an F/C-style sparsity pattern (e.g,
      F/C pattern generates a P0)?
\end{enumerate}

\input{commLayer}

\end{document}
